Thesis: draft

### 2021-03-13 draft.tex
continued writing stubbornness section

\subsection{The majority model}

@@

It is well known that in real-life, you can find people that are really persistent towards their own opinion and that are not open to opinions of others. They are convinced that they carry the truth with them and that the other people, with different world-views, are not capable of giving them any new and useful insights. They are close-minded in contrast to people with an open-mind that are willing to learn from others and to reconsider their beliefs. Of course this is put a bit simplistic and there are many shades of gray between this black and white statement. However, there are groups in society, often with more extreme opinions, that are know to be really inflexible towards changing that (extreme) opinion (note here that being stubborn and being extreme doesn't need to go hand in hand, you can obviously be stubborn without being extreme and vice versa; however, it is often widely accepted that there is a tendency that people with extreme opinions are more persistent towards these opinions than people with more moderate beliefs; this idea of correlating extreme people with a more stubborn character is also often used in models, see for example \cite{Deffuant2006}\cite{Turner}).\\ 
Knowing the impact of such stubborn groups on the opinions of others is of great importance, especially with the rise of social media which makes the reach of such extreme and persistent groups a lot bigger. Before the era of social media, one could maybe argue that these extreme and stubborn groups are often isolated and therefore wouldn't reach too many people. So their impact could only be of a minor nature (note that this is only a statement that could have been given, there are of course plenty of examples throughout human history that contradict this argument). Today, this possible argument is no longer valid; social media enables the contact with all sorts of beliefs and all sorts of people with almost no constraints. Extreme and stubborn groups that may have been isolated before no longer need to be so and this may have huge impacts. \\
What is the impact of having stubborn/extreme people in your society? Does it matter if the stubborn people are grouped together or if they are distributed evenly throughout your system? Are there ways to reduce the possible impacts of stubbornness and/or extremism? Can these stubborn groups lead to polarization or even overtake the opinions in the system? To answer these questions, several models have been proposed to include some kind of stubbornness/resistance in the system under study.

@@

\subsection{Stubbornness in binary models}
% not finished!!
Playing around with the threshold $\epsilon_i$ of the agents to include a notion of stubbornness, such as one can do for the bounded confidence models, is obviously not possible for binary/discrete models. There are, however, many other ways to include some kind of stubborn people in these models. In the following, two different possibilities will be discussed. One small remark must be made here: previously, the link between stubbornness and extremism was often made and it was often assumed that the more extreme somebodies opinions are, the more persistent that somebody is towards those opinions. In binary models, however, there are only two possible opinions that can be taken on and in that sense it is less useful to talk about extreme and moderate opinions. However, we can still have people that rather keep their own opinion than be convinced by friends to switch to the other opinion.\\
\newline
\textbf{Including a stubbornness parameter}\\
\newline
One way of creating those stubborn people in the binary model is by giving each node $i$ a stubbornness/resistance parameter $r_i$. This resistance parameters can have values between 0 and 1, $0 \leqslant r_i \leqslant 1$. 
\begin{itemize}
	\item If $r_i = 0$, the node is not stubborn and can be easily convinced of the value of the other opinion (according to which opinion dynamics model that is implemented: majority, probabilistic majority,...)
	\item If $r_i = 1$, the node is completely stubborn and there is no change of convincing him/here of the other idea
	\item If $0 < r_i < 1$, the node has some resistance but can still be convinced to switch to the other opinion
\end{itemize}
The resistance parameter $r_i$ can be different for all nodes or can be the same. One can also play around with the fraction of nodes that is stubborn, for example, there can be a fraction of nodes that is completely stubborn ($r = 1$) and a remaining fraction that is not stubborn at all ($r = 0$). Many other possibilities can be found according to how the stubborn people are distributed and according to how stubborn they are made. This makes the model very flexible.\\
\newline
\textbf{The threshold model}
\newline
% NOT FINISHED!!

Note that the above examples are only a handful of the possible ways to include some notion of stubbornness in the model under study.

>>>>

\subsection{The majority model}\label{majModel}

@@

It is well known that in real-life, you can find people that are really persistent towards their own opinion and that are not open to opinions of others. They are convinced that they carry the truth with them and that the other people, with different world-views, are not capable of giving them any new and useful insights. They are close-minded in contrast to people with an open-mind that are willing to learn from others and to reconsider their beliefs. Of course this is put a bit simplistic and there are many shades of gray between this black and white statement. However, there are groups in society, often with more extreme opinions, that are know to be really inflexible towards changing that (extreme) opinion (note here that being stubborn and being extreme doesn't need to go hand in hand, you can obviously be stubborn without being extreme and vice versa; however, it is often widely accepted that there is a tendency that people with extreme opinions are more persistent towards these opinions than people with more moderate beliefs; this idea of correlating extreme people with a more stubborn character is also often used in models, see for example \cite{Deffuant2006}\cite{Turner}).\\ 
Knowing the impact of such stubborn groups on the opinions of others is of great importance, especially with the rise of social media which makes the reach of such extreme and persistent groups a lot bigger. Before the era of social media, one could maybe argue that these extreme and stubborn groups are often isolated and therefore wouldn't reach too many people. So their impact could only be of a minor nature (note that this is only a statement that could have been given, there are of course plenty of examples throughout human history that contradict this argument; to only mention a few: Nazism in the 1930s and 1940s, the Ku Klux Klan in the late 19th and begin 20th century, the rise of communism in the 20th century, etc). Today, this possible argument is no longer valid; social media enables the contact with all sorts of beliefs and all sorts of people with almost no constraints. Extreme and stubborn groups that may have been isolated before no longer need to be so and this may have huge impacts. \\
What is the impact of having stubborn/extreme people in your society? Does it matter if the stubborn people are grouped together or if they are distributed evenly throughout your system? Are there ways to reduce the possible impacts of stubbornness and/or extremism? Can these stubborn groups lead to polarization or even overtake the opinions in the system? To answer these questions, several models have been proposed to include some kind of stubbornness/resistance in the system under study.

@@

\subsection{Stubbornness in binary models}
% not finished!!
Playing around with the threshold $\epsilon_i$ of the agents to include a notion of stubbornness, such as one can do for the bounded confidence models, is obviously not possible for binary/discrete models. There are, however, many other ways to include some kind of stubborn people in these models. In the following, two different possibilities will be discussed. One small remark must be made here: previously, the link between stubbornness and extremism was often made and it was often assumed that the more extreme somebodies opinions are, the more persistent that somebody is towards those opinions. In binary models, however, there are only two possible opinions that can be taken on and in that sense it is less useful to talk about extreme and moderate opinions. However, we can still have people that rather keep their own opinion than be convinced by friends to switch to the other opinion.\\
\textbf{Including a stubbornness parameter}\\
One way of creating those stubborn people in the binary model is by giving each node $i$ a stubbornness/resistance parameter $r_i$. This resistance parameter can have values between 0 and 1, $0 \leqslant r_i \leqslant 1$. 
	\item If $r_i = 0$, the node is not stubborn and can be easily convinced of the value of the other opinion (according to which opinion dynamics model that is implemented: majority, probabilistic majority,...)
	\item If $r_i = 1$, the node is completely stubborn and there is no change of convincing him/here of the other idea
	\item If $0 < r_i < 1$, the node has some resistance but can still be convinced to switch to the other opinion, even though he/she will be less likely to do so 
The resistance parameter $r_i$ can be different or the same for all nodes. One can also play around with the fraction of nodes that is stubborn, for example, there can be a fraction of nodes that is completely stubborn ($r = 1$) and a remaining fraction that is not stubborn at all ($r = 0$). Many other possibilities can be found according to how the stubborn people are distributed and according to how stubborn they are made. This makes the model very flexible.\\
\textbf{The threshold model}
In the threshold model one introduces a notion of stubbornness by introducing a threshold parameter $T$. This parameter can be either the same for all the nodes ($T$) or can be made individual ($T_i$). The main idea is then that nodes can only change their opinion to the other opinion if the fraction of friends with that opinion ($p$) is bigger than $T$ ($p > T$). The idea behind this threshold model is that people are a bit hesitant to change their opinion, but if they have enough friends with another opinion they might be convinced of that opinion as well. The more persistent you are, the more friends you need in order to be convinced of another idea. \\
There are different possibilities to implement this kind of model. One way, which makes use of the majority model, is as follows. First the fraction of friends with opinion A and opinion B, $p_A$ and $p_B$ respectively, is determined; then the following rules are used to update the opinion
	\item If $p_A > p_B$ and $p_A > T$, opinion A is adopted
	\item If $p_B > p_A$ and $p_B > T$, opinion B is adopted
	\item In all other cases, the current opinion is not changed
If $T = 0$, this model reduces to the pure majority model described in Section \ref{majModel}. If $T = 1$, the nodes don't change their opinion. For $0 < T < 1$, the model is a majority model, but the nodes are a bit more persistent towards their own opinion and need a larger fraction of friends to convince them to switch to the other viewpoint.\\
% Talk about probabilistic implementation?
Note that the above examples are only a handful of the possible ways to include some notion of stubbornness in the model under study, many more exist or can be invented.
